---
title: "p8105_hw6_yc3577"
author: "Yimeng Cai"
date: "11/30/2023"
output: github_document
---
```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(broom)
library(modelr)
library(mgcv)
library(purrr)
library(dplyr)
library(stats)
```
Problem 1


### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)

homicide_df
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)
baltimore_glm

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
baltimore_glm
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)

model_results
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

model_results
```

Problem 2
For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

First, we need to load in the data. 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```
The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:
- r̂^ 2
- log(β̂ 1∗β̂ 2)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 
```{r}
bootstrap_sample = function(data) {
  sample_frac(data, replace = TRUE)
}


bootstrap_tibble = 
  tibble(number = 1:5000) |>
  mutate(
    sample = map(number, \(i) bootstrap_sample(weather_df)))

bootstrap_tibble

bootstrap_results1 = bootstrap_tibble |>
  mutate(
    models = map(sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    r_squared = map_dbl(models, \(model) summary(model)$r.squared)) |> 
  select(-models) |> 
  unnest(results) 

bootstrap_results1

bootstrap_results2 =
  bootstrap_results1 |>
  group_by(term) |>
  select(number, estimate, r_squared, term) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |>
  mutate(
    log_beta = log(tmin * prcp)
  )

bootstrap_results2
  
```
Plot the distribution of your estimates, and describe these in words.

```{r}
# Then we need to draw the distribution of log_beta

beta_dist =
  bootstrap_results2 |>
  ggplot(aes(x = log_beta)) +
  geom_density(color = 'pink') +
  theme_minimal() +
  labs(
    x = 'Log Beta estimate',
    y = 'Frequency',
    title = 'Distribution of Log beta'
  )

beta_dist

hist(bootstrap_results2$log_beta, main = "Distribution of Log beta", xlab = "Log beta", col = "pink")
```
# The distribution of Log_beta is above with Pink histogram and pink density line. It is unimodel. From two graphs, we can see they are left skewed distributed with a long tail on the left. The range is from -14 to -4 and the center of the mode is around -6. This suggests that there are a few samples being strapped having a small product of beta coefficients. 




```{r}
# Then we need to draw the distribution of r^2

rsquared_dist =
  bootstrap_results2 |>
  ggplot(aes(x = r_squared)) +
  geom_density(color = 'skyblue') +
  theme_minimal() +
  labs(
    x = 'R Squared estimate',
    y = 'Frequency',
    title = 'Distribution of R Squared'
  )

rsquared_dist

hist(bootstrap_results2$r_squared, main = "Distribution of R-squared", xlab = "R-squared", col = "skyblue")
```
# The distribution of R_squared is above with blue histogram and blue density curve. It is unimodel. From two graphs, we can see they are also left skewed a little with a small tail on the left. But the distribution is closer to normal than the log_beta graph. The range is from 0.86 to 0.95 and the center of the mode is around 0.92.

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 1∗β̂ 2).

```{r}
# Then we can construct the 95% confidence interval for rsquared
rsquared_quantile = 
  bootstrap_results2 |>
  drop_na()|>
  summarize(
    CI_Low = quantile(r_squared, 0.025),
    CI_High = quantile(r_squared, 0.975)
  )
rsquared_quantile
```
# From calculation, the 95% confidence interval for r_squared is (0.8828948, 0.9273118) which means that we have 95 out of 100 times the estimate r_squared will fall between the interval of (0.8828948, 0.9273118).

```{r}
# Then we can construct the 95% confidence interval for logbeta
logbeta_quantile = 
  bootstrap_results2 |>
  drop_na()|>
  summarize(
    CI_Low = quantile(log_beta, 0.025),
    CI_High = quantile(log_beta, 0.975)
  )
logbeta_quantile
```
# From calculation, the 95% confidence interval for log_(betatmin * betaprcp) is (-8.998584, -4.589433) which means that we have 95 out of 100 times the estimate log_(betatmin * betaprcp) will fall between the interval of (-8.998584, -4.589433).


Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. 
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
birthweight = 
  read_csv('birthweight.csv') |>
  janitor::clean_names() |>
  mutate(
    babysex = recode(babysex, '1' = 'male', '2' = 'female'),
    frace = recode(frace, '1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Puerto Rican', '8' = 'Other', '9' = 'Unknown'),
    mrace = recode(mrace, '1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Puerto Rican', '8' = 'Other'),
    malform = recode(malform, '0' = 'absent', '1' = 'present')
  ) |>
  drop_na()

birthweight
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. 
```{r}
birthweight_model = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks +
                          malform + menarche + mheight + momage + mrace + parity + pnumlbw +
                          pnumsga + ppbmi + ppwt + smoken + wtgain, data = birthweight)
birthweight_model
```


Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
birthweight_df = 
  birthweight |>
  mutate(
    fitted_values = as.vector(fitted(birthweight_model)),
    residuals = as.vector(residuals(birthweight_model)))

birthweight_df

ggplot(birthweight_df, aes(x = fitted_values, y = residuals)) +
  geom_point(aes(color = "Residuals"), alpha = 0.5) +
  geom_line(stat = "smooth", method = "lm", se = FALSE, alpha = 1, color = "blue") +
  theme_minimal() +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals")

```


Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
```{r}
# Model with head circumference, length, sex, and all interactions
model1 <- lm(bwt ~ blength + gaweeks, data = birthweight)
model1
```
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
# Model with head circumference, length, sex, and all interactions
model2 <- lm(bwt ~ bhead + blength + babysex + bhead * blength * babysex, data = birthweight)
model2
```

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}
set.seed(12)

train_df = sample_n(birthweight, 80)
train_df
test_df = anti_join(birthweight, train_df)
test_df


rmse_bw_train = rmse(birthweight_model, train_df)
rmse_bw_test = rmse(birthweight_model, test_df)
rmse_1_train = rmse(model1, train_df)
rmse_1_test = rmse(model1, test_df)
rmse_2_train = rmse(model2, train_df)
rmse_2_test = rmse(model2, test_df)
```


Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

