---
title: "p8105_hw6_yc3577"
author: "Yimeng Cai"
date: "11/30/2023"
output: html_document
---
```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(broom)
library(modelr)
library(mgcv)
library(purrr)
library(dplyr)
library(stats)
```


Problem 2
For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

First, we need to load in the data. 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```
The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:
- r̂^ 2
- log(β̂ 1∗β̂ 2)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 
```{r}
bootstrap_sample = function(data) {
  sample_frac(data, replace = TRUE)
}


bootstrap_results1 = 
  tibble(number = 1:5000) |>
  mutate(
    sample = map(number, \(i) bootstrap_sample(weather_df)),
    models = map(sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    r_squared = map_dbl(models, \(model) summary(model)$r.squared)) |> 
  select(-models) |> 
  unnest(results) 

bootstrap_results1

bootstrap_results2 =
  bootstrap_results1 |>
  group_by(term) |>
  select(number, term, estimate, r_squared) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |>
  mutate(
    log_beta = log(tmin * prcp)
  )

bootstrap_results2
  
```
Plot the distribution of your estimates, and describe these in words.

```{r}
# Then we need to draw the distribution of log_beta

beta_dist =
  bootstrap_results2 |>
  ggplot(aes(x = log_beta)) +
  geom_density(color = 'pink') +
  theme_minimal() +
  labs(
    x = 'Log Beta estimate',
    y = 'Strap Number',
    title = 'Distribution of Log beta'
  )

beta_dist

hist(bootstrap_results2$log_beta, main = "Distribution of Log beta", xlab = "Log beta", col = "pink")
```





```{r}
# Then we need to draw the distribution of r^2

rsquared_dist =
  bootstrap_results2 |>
  ggplot(aes(x = r_squared)) +
  geom_density(color = 'skyblue') +
  theme_minimal() +
  labs(
    x = 'R Squared estimate',
    y = 'Strap Number',
    title = 'Distribution of R Squared'
  )

rsquared_dist

hist(bootstrap_results2$r_squared, main = "Distribution of R-squared", xlab = "R-squared", col = "skyblue")
```

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 1∗β̂ 2).

```{r}
# Then we can construct the 95% confidence interval
rsquared_quantile = 
  bootstrap_results2 |>
  drop_na()|>
  summarize(
    CI_Low = quantile(r_squared, 0.025),
    CI_High = quantile(r_squared, 0.975)
  )
rsquared_quantile
```

```{r}
# Then we can construct the 95% confidence interval
logbeta_quantile = 
  bootstrap_results2 |>
  drop_na()|>
  summarize(
    CI_Low = quantile(log_beta, 0.025),
    CI_High = quantile(log_beta, 0.975)
  )
logbeta_quantile
```
Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. 
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
birthweight = 
  read_csv('birthweight.csv') |>
  janitor::clean_names() |>
  mutate(
    babysex = recode(babysex, '1' = 'male', '2' = 'female'),
    frace = recode(frace, '1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Puerto Rican', '8' = 'Other', '9' = 'Unknown'),
    mrace = recode(mrace, '1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Puerto Rican', '8' = 'Other'),
    malform = recode(malform, '0' = 'absent', '1' = 'present')
  ) |>
  drop_na()

birthweight
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. 
```{r}
birthweight_model = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks +
                          malform + menarche + mheight + momage + mrace + parity + pnumlbw +
                          pnumsga + ppbmi + ppwt + smoken + wtgain, data = birthweight)
birthweight_model
```


Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
birthweight_df = 
  birthweight |>
  mutate(
    fitted_values = as.vector(fitted(birthweight_model)),
    residuals = as.vector(residuals(birthweight_model)))

birthweight_df

ggplot(birthweight_df, aes(x = fitted_values, y = residuals)) +
  geom_point(aes(color = "Residuals"), alpha = 0.5) +
  geom_line(stat = "smooth", method = "lm", se = FALSE, alpha = 1, color = "blue") +
  theme_minimal() +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals")

```


Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
```{r}
# Model with head circumference, length, sex, and all interactions
model1 <- lm(bwt ~ blength + gaweeks, data = birthweight)
model1
```
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
# Model with head circumference, length, sex, and all interactions
model2 <- lm(bwt ~ bhead + blength + babysex + bhead * blength * babysex, data = birthweight)
model2
```

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}
set.seed(12)

train_df = sample_n(birthweight, 80)
train_df
test_df = anti_join(birthweight, train_df)
test_df


rmse_bw_train = rmse(birthweight_model, train_df)
rmse_bw_test = rmse(birthweight_model, test_df)
rmse_1_train = rmse(model1, train_df)
rmse_1_test = rmse(model1, test_df)
rmse_2_train = rmse(model2, train_df)
rmse_2_test = rmse(model2, test_df)
```


Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

